Outbreak Threshold Detection

### 2023-03-03 notebooks/Julia/simulation-detection.qmd
Add Tau-leaping Add RegularJump tau-leaping implementation. Needed to include N explicitly in the states because calculating rates was causing an issue using u[3] i.e. Z state. Need to look into why this is an issue, but for now, tau-leaping produces the expected result

```{julia}
R₀ = 16
γ = 1/8
μ = 1 / (365 * 62.5)
β = R₀ * (γ + μ)
p = (; β, γ, μ)
```

```{julia}
tmax = 250
δt = 0.001
tspan = (0.0, tmax)
tlength = length(tspan[1]:δt:tspan[2])
Random.seed!(1234);
```

```{julia}
infection_rate(u, p, t) = p.β * u[1] * u[2] / (u[1] + u[2] + u[3])      # β*X*Y / N

@@

```{julia}
recovery_rate(u, p, t) = p.γ * u[2]     # γ*Y

@@

```{julia}
birth_rate(u, p, t) = p.μ * (u[1] + u[2] + u[3])        # μ * N

@@

```{julia}
X_death_rate(u, p, t) = p.μ * u[1]       # μ * X

@@

```{julia}
Y_death_rate(u, p, t) = p.μ * u[2]       # μ * Y

@@

```{julia}
Z_death_rate(u, p, t) = p.μ * u[3]       # μ * R

@@

```{julia}
prob = DiscreteProblem(u₀, tspan, p)
jump_prob = JumpProblem(prob, Diredt(), jumps...)
sol = solve(jump_prob, SSAStepper())

sol_array = zeros(4, length(sol))
sol_array[1:3, :] = Array(sol)
sol_array[4, :] = sum(sol_array[1:3, :], dims = 1);
```
```{julia}
f = Figure()
ax = Axis(
    f[1, 1],
    xlabel = "Time",
    ylabel = "Number"
)

>>>>

```{julia}
R₀ = 16
γ = 1/8
μ = 1 / (365 * 62.5)
β = R₀ * (γ + μ)
p = (β, γ, μ)
```

```{julia}
tmax = 250.0
δt = 0.001
tspan = (0.0, tmax)
tlength = length(tspan[1]:δt:tspan[2])
Random.seed!(1234);
```

```{julia}
function rate(out, u, p, t)
    out[1] = p[1] * u[1] * u[2] / u[4]      # β * X * Y / N
    out[2] = p[2] * u[2]                    # ν * Y
    out[3] = p[3] * u[4]                    # μ * N
    out[4] = p[3] * u[1]                    # μ * X
    out[5] = p[3] * u[2]                    # μ * Y
    out[6] = p[3] * u[3]                    # μ * Z
    nothing
end
```

```{julia}
c = zeros(4, 6)
# X + Y --> Y
c[1, 1] = -1        # X -> X - 1
c[2, 1] = 1         # Y -> Y + 1

# Y --> Z
c[2, 2] = -1        # Y -> Y - 1
c[3, 2] = 1         # Z -> Z + 1

# X --> 2X
c[1, 3] = 1         # X -> X + 1
c[4, 3] = 1         # N -> N + 1

# X --> 0X
c[1, 4] = -1        # X -> X - 1
c[4, 4] = -1        # N -> N - 1

# Y --> 0Y
c[2, 5] = -1        # Y -> Y - 1
c[4, 5] = -1        # N -> N - 1

# Z --> 0Z
c[3, 6] = -1        # Z -> Z - 1
c[4, 6] = -1        # N -> N - 1

function change!(du, u, p, t, counts, mark)
    mul!(du, c, counts)
    nothing
end
```

```{julia}
rj = RegularJump(rate, change!, 6)
```

```{julia}
u₀ = [800, 20, 800, NaN]
u₀[4] = sum(u₀[1:3])
R₀ = 16
γ = 1/8
μ = 20 / (365 * 62.5)
β = R₀ * (γ + μ)
p = (β, γ, μ)
prob = DiscreteProblem(u₀, tspan, p)
jump_prob = JumpProblem(prob, Direct(), rj)
sol = solve(jump_prob, TauLeaping(); dt = δt)
sol_array = zeros(5, length(sol))
sol_array[1:4, :] = Array(sol)
sol_array[5, :] = sum(sol_array[1:3, :], dims = 1);

sol_array[4, :] == sol_array[5, :]
```

```{julia}
f = Figure()
ax = Axis(
    f[1, 1],
    xlabel = "Time",
    ylabel = "Number"
)

lines!(ax, sol.t, sol_array[1, :], linewidth = 3, label = "X(t)")
lines!(ax, sol.t, sol_array[2, :], linewidth = 3, label = "Y(t)")
lines!(ax, sol.t, sol_array[3, :], linewidth = 3, label = "Z(t)")
lines!(ax, sol.t, sol_array[4, :], linewidth = 3, label = "N(t) jump")
lines!(ax, sol.t, sol_array[5, :], linewidth = 3, label = "N(t) calc")

```{julia}
infection_rate(u, p, t) = p[1] * u[1] * u[2] / (u[1] + u[2] + u[3])      # β*X*Y / N

@@

```{julia}
recovery_rate(u, p, t) = p[2] * u[2]     # γ*Y

@@

```{julia}
birth_rate(u, p, t) = p[3] * (u[1] + u[2] + u[3])        # μ * N

@@

```{julia}
X_death_rate(u, p, t) = p[3] * u[1]       # μ * X

@@

```{julia}
Y_death_rate(u, p, t) = p[3] * u[2]       # μ * Y

@@

```{julia}
Z_death_rate(u, p, t) = p[3] * u[3]       # μ * R

@@

```{julia}
prob = DiscreteProblem(u₀, tspan, p)
jump_prob = JumpProblem(prob, Direct(), jumps...)
sol = solve(jump_prob, SSAStepper())

sol_array = zeros(4, length(sol))
sol_array[1:4, :] = Array(sol)
sol_array[4, :] = sum(sol_array[1:3, :], dims = 1);
```
```{julia}
f = Figure()
ax = Axis(
    f[1, 1],
    xlabel = "Time",
    ylabel = "Number"
)
